A hazards approach to the biometric analysis of infant mortality

### 2023-03-06 paper2.Rmd
first commit

@@

```{r first, fig.cap = "The periods 1861--1890 and 1921--1950 in Västerbotten, Bourgeois-Pichat method."}
source("R/drawBP.R")
oldpar <- par(mfrow = c(1,2))
##
r1 <- with(ginfant[ginfant$period == "1861-1890", ], eha::Surv(enter, exit, event))
r1 <- eha::risksets(r1, members = FALSE)
drawBP(r1, main = "1861-1890", ylim = c(0, 0.1))
##
##r2 <- with(ginfant[ginfant$period == "1891-1920", ], eha::Surv(enter, exit, event))
##r2 <- eha::risksets(r2, members = FALSE)
##drawBP(r2, main = "1891-1920", ylim = c(0, 0.1))
##
r3 <- with(ginfant[ginfant$period == "1921-1950", ], eha::Surv(enter, exit, event))
r3 <- eha::risksets(r3, members = FALSE)
drawBP(r3, main = "1921-1950", ylim = c(0, 0.1))
par(oldpar)
```

@@

```{r second, fig.cap = "The periods 1861--1890 and 1921--1950 in Västerbotten, hazards method."}
source("R/drawBB.R")
oldpar <- par(mfrow = c(1,2))
##
##r1 <- with(ginfant[ginfant$period == "1861-1890", ], eha::Surv(enter, exit, event))
##r1 <- eha::risksets(r1, members = FALSE)
drawBB(r1, main = "1861-1890", ylim = c(0, 0.115))
##
##r2 <- with(ginfant[ginfant$period == "1891-1920", ], eha::Surv(enter, exit, event))
##r2 <- eha::risksets(r2, members = FALSE)
##drawBP(r2, main = "1891-1920", ylim = c(0, 0.1))
##
##r3 <- with(ginfant[ginfant$period == "1921-1950", ], eha::Surv(enter, exit, event))
##r3 <- eha::risksets(r3, members = FALSE)
drawBB(r3, main = "1921-1950", ylim = c(0, 0.115))
par(oldpar)
```

@@

```{r plotfun}
pf <- function(x){
   ## x is a dataframe on g scale
   rs <- with(x, eha::risksets(Surv(enter, exit, event), members = FALSE))
   oldpar <- par(mfrow = c(1, 2))
   drawBB(rs, main = "Hazards method")
   drawBP(rs, main = "B-P method")
##
par(oldpar)
}
```

@@

```{r lowumee, fig.cap = "Västerbotten 1921--1950."}
##source("R/compare.R")
##xx <- bio(ume, timescale = "day", graph = TRUE)
##post <- age.window(ume, c(g(28), g(365)))
##rs.post <- risksets()
ume <- ginfant[ginfant$birthdate > "1920-12-31", ]
pf(ume)
```

@@

```{r bioume9120, fig.cap = "Västerbotten 1891--1920."}
##source("R/compare.R")
##xx <- bio(ume, timescale = "day", graph = TRUE)
##post <- age.window(ume, c(g(28), g(365)))
##rs.post <- risksets()
ume$exit <- g(ume$exit)
ume$enter <- g(ume$enter)
pf(ume)
```

@@

```{r bioume6190, fig.cap = "Västerbotten 1861--1890."}
##source("R/compare.R")
##xx <- bio(ume, timescale = "day", graph = TRUE)
##post <- age.window(ume, c(g(28), g(365)))
##rs.post <- risksets()
ume$exit <- g(ume$exit)
ume$enter <- g(ume$enter)
pf(ume)
```

@@

```{r scahigh, fig.cap = "Scania 1710--1800."}
sca1 <- readRDS("Rdata/sca1.rds")
sca1 <- sca1[sca1$BirthDate < "1801-01-01", ]
sca1$exit <- g(sca1$exit)
sca1$enter <- g(sca1$enter)
rs <- with(sca1, risksets(Surv(enter, exit, event), members = FALSE))
oldpar <- par(mfrow = c(1, 2))
drawBB(rs, main = "Hazards method")
drawBP(rs, main = "B-P method")
par(oldpar)
```

@@

```{r skum, fig.cap = "Skellefteå 1821--1838."}
library(skum)
births <- obs[obs$starttyp == 2 &
                 obs$foddat > as.Date("1820-12-31") &
                 obs$foddat <= as.Date("1838-12-31"), ]
births$event <- births$sluttyp == 2
births$enter <- 365 * births$enter
births$exit <- 365 * births$exit
addup <- births$exit <= 0
births$exit[addup] <- 0.25
infa <- age.window(births, c(0, 365))
infa$enter <- g(infa$enter)
infa$exit <- g(infa$exit)
pf(infa)
```

>>>>

In this version the cumulative hazard functions are depicted as-is, but on the
log-cube time scale. See the new **R** function *drawBB2*.

@@

```{r first, fig.cap = "The periods 1861--1890 and 1921--1950 in Västerbotten, Bourgeois-Pichat method."}
source("R/drawBP.R")
oldpar <- par(mfrow = c(1,2))
##
r1 <- with(ginfant[ginfant$period == "1861-1890", ], eha::Surv(enter, exit, event))
g1 <- infant[infant$period == "1861-1890", ]
r1 <- eha::risksets(r1, members = FALSE)
drawBP(r1, main = "1861-1890", ylim = c(0, 0.1))
##
##r2 <- with(ginfant[ginfant$period == "1891-1920", ], eha::Surv(enter, exit, event))
##r2 <- eha::risksets(r2, members = FALSE)
##drawBP(r2, main = "1891-1920", ylim = c(0, 0.1))
##
g3 <- infant[infant$period == "1921-1950", ]
r3 <- with(ginfant[ginfant$period == "1921-1950", ], eha::Surv(enter, exit, event))
r3 <- eha::risksets(r3, members = FALSE)
drawBP(r3, main = "1921-1950", ylim = c(0, 0.1))
par(oldpar)
```

@@

```{r second, fig.cap = "The periods 1861--1890 and 1921--1950 in Västerbotten, hazards method."}
source("R/drawBB2.R")
oldpar <- par(mfrow = c(1,2))
##
##r1 <- with(ginfant[ginfant$period == "1861-1890", ], eha::Surv(enter, exit, event))
##r1 <- eha::risksets(r1, members = FALSE)
drawBB2(g1, main = "1861-1890")
##
##r2 <- with(ginfant[ginfant$period == "1891-1920", ], eha::Surv(enter, exit, event))
##r2 <- eha::risksets(r2, members = FALSE)
##drawBP(r2, main = "1891-1920", ylim = c(0, 0.1))
##
##r3 <- with(ginfant[ginfant$period == "1921-1950", ], eha::Surv(enter, exit, event))
##r3 <- eha::risksets(r3, members = FALSE)
drawBB2(g3, main = "1921-1950")
par(oldpar)
```

@@

```{r plotfun}
pf <- function(x){
   ## x is a dataframe NOT on g scale
    y <- x
    y$enter <- g(y$enter)
    y$exit <- g(y$exit)
   rs <- with(y, eha::risksets(Surv(enter, exit, event), members = FALSE))
   oldpar <- par(mfrow = c(1, 2))
   drawBB2(x, main = "Hazards method")
   drawBP(rs, main = "B-P method")
##
par(oldpar)
}
```

@@

```{r lowumee, fig.cap = "Västerbotten 1921--1950."}
##source("R/compare.R")
##xx <- bio(ume, timescale = "day", graph = TRUE)
##post <- age.window(ume, c(g(28), g(365)))
##rs.post <- risksets()
ume <- infant[infant$birthdate > "1920-12-31", ]
pf(ume)
```

@@

```{r bioume9120, fig.cap = "Västerbotten 1891--1920."}
##source("R/compare.R")
##xx <- bio(ume, timescale = "day", graph = TRUE)
##post <- age.window(ume, c(g(28), g(365)))
##rs.post <- risksets()
##ume$exit <- g(ume$exit)
##ume$enter <- g(ume$enter)
pf(ume)
```

@@

```{r bioume6190, fig.cap = "Västerbotten 1861--1890."}
##source("R/compare.R")
##xx <- bio(ume, timescale = "day", graph = TRUE)
##post <- age.window(ume, c(g(28), g(365)))
##rs.post <- risksets()
##ume$exit <- g(ume$exit)
##ume$enter <- g(ume$enter)
pf(ume)
```

@@

```{r scahigh, fig.cap = "Scania 1710--1800."}
sca1 <- readRDS("Rdata/sca1.rds")
sca1 <- sca1[sca1$BirthDate < "1801-01-01", ]
sca11 <- sca1
sca1$exit <- g(sca1$exit)
sca1$enter <- g(sca1$enter)
rs <- with(sca1, risksets(Surv(enter, exit, event), members = FALSE))
oldpar <- par(mfrow = c(1, 2))
drawBB2(sca11, main = "Hazards method")
drawBP(rs, main = "B-P method")
par(oldpar)
```

@@

```{r skum, fig.cap = "Skellefteå 1821--1838."}
library(skum)
births <- obs[obs$starttyp == 2 &
                 obs$foddat > as.Date("1820-12-31") &
                 obs$foddat <= as.Date("1838-12-31"), ]
births$event <- births$sluttyp == 2
births$enter <- 365 * births$enter
births$exit <- 365 * births$exit
addup <- births$exit <= 0
births$exit[addup] <- 0.25
infa <- age.window(births, c(0, 365))
##infa$enter <- g(infa$enter)
##infa$exit <- g(infa$exit)
pf(infa)
```
